---
title: Applied Exercises for Chapter 2
author: Jean Cochrane
date: 8 July 2018
---

### Exercise 8

a. First, let's read the `college.csv` data into R:

```{r}
college = read.csv('data/college.csv')
```

b. Next, we can change the rownames so that the rows are labelled with the
college name but that name is not considered part of the data:

```{r}
# Label each row with the name of the college
rownames(college) = college[,1]

# Remove the name of the college (column 1) from the data
college = college[,-1]
```

c.i. Numerical summary of the dataset:

```{r}
summary(college)
```

c.ii. A scatterplot matrix of the first ten columns of the data:

```{r}
pairs(college[,1:10])
```

c.iii. A boxplot of `Outstate` vs. `Private`:

```{r}
outstate = college$Outstate
private = as.factor(college$Private)

plot(private, outstate, xlab="Private", ylab="Out of State Tuition")
```

c.iv. Create a new qualitative variable, `Elite`, by binning `Top10perc`:

```{r}
Elite = rep("No", nrow(college))
Elite[college$Top10perc>50] = "Yes"
Elite = as.factor(Elite)

college = data.frame(college, Elite)
summary(college$Elite)
```

So there are 78 "elite" universities in this dataset.

Side-by-side boxplots of `Outstate` versus `Elite`:

```{r}
plot(Elite, outstate, xlab="Elite", ylab="Out of State Tuition")
```

c.v. Histograms of some of the quantitative variables:

```{r}
par(mfrow=c(2,2))
hist(college$Room.Board, plot=TRUE)
hist(college$Books, plot=TRUE)
hist(college$Personal, plot=TRUE)
hist(college$Expend, plot=TRUE)
```
