---
title: Applied Exercises for Chapter 2
author: Jean Cochrane
date: 8 July 2018
---

# 3.6.1 Libraries

Load R packages:

```{r}
library(MASS)
library(ISLR)
```

# 3.6.2 Simple Linear Regression

Print column names:

```{r}
names(Boston)
```

Regress `medv` on `lstat` (median home value on percent of households with low
socioeconomic status):

```{r}
lm.fit = lm(medv~lstat, data=Boston)
```

Print a summary of the model:

```{r}
summary(lm.fit)
```

Get coefficients of the model:

```{r}
coef(lm.fit)
```

Get confidence interval of the model coefficients:

```{r}
confint(lm.fit)
```

Predict values for `medv` given values of `lstat`, returning
a confidence interval:

```{r}
predict(lm.fit, data.frame(lstat=c(5, 10, 15)), interval="confidence")
```

Now returning a prediction interval:

```{r}
predict(lm.fit, data.frame(lstat=c(5, 10, 15)), interval="prediction")
```

Plot the least squares regression line:

```{r}
plot(Boston$lstat, Boston$medv)
abline(lm.fit, lwd=3, col="red")
```

Experiment with plotting options:

```{r}
plot(1:20, 1:20, pch=1:20)
```

Generate default diagnostic plots for the model:

```{r}
par(mfrow=c(2, 2))
plot(lm.fit)
```
