---
title: Applied Exercises for Chapter 2
author: Jean Cochrane
date: 8 July 2018
---

# 3.6.1 Libraries

Load R packages:

```{r}
library(MASS)
library(ISLR)
```

# 3.6.2 Simple Linear Regression

Print column names:

```{r}
names(Boston)
```

Regress `medv` on `lstat` (median home value on percent of households with low
socioeconomic status):

```{r}
lm.fit = lm(medv~lstat, data=Boston)
```

Print a summary of the model:

```{r}
summary(lm.fit)
```

Get coefficients of the model:

```{r}
coef(lm.fit)
```

Get confidence interval of the model coefficients:

```{r}
confint(lm.fit)
```

Predict values for `medv` given values of `lstat`, returning
a confidence interval:

```{r}
predict(lm.fit, data.frame(lstat=c(5, 10, 15)), interval="confidence")
```

Now returning a prediction interval:

```{r}
predict(lm.fit, data.frame(lstat=c(5, 10, 15)), interval="prediction")
```

Plot the least squares regression line:

```{r}
plot(Boston$lstat, Boston$medv)
abline(lm.fit, lwd=3, col="red")
```

Experiment with plotting options:

```{r}
plot(1:20, 1:20, pch=1:20)
```

Generate default diagnostic plots for the model:

```{r}
par(mfrow=c(2, 2))
plot(lm.fit)
```

Plot the residuals:

```{r}
plot(predict(lm.fit), residuals(lm.fit))
```

Plot the studentized residuals:

```{r}
plot(predict(lm.fit), rstudent(lm.fit))
```

Looks like there's some nonlinearity!

Compute leverage statistics:

```{r}
plot(hatvalues(lm.fit))
```

Find the index of the observation with the largest leverage statistic:

```{r}
which.max(hatvalues(lm.fit))
```

And the observation itself:

```{r}
Boston[which.max(hatvalues(lm.fit)), 'lstat']
```

# 3.6.3 Multiple Linear Regression

To fit a model with `n` predictors, use the following syntax:

```
lm.fit = lm(y~x1+x2+...+x3)
```

Fit a model to `medv` using `lstat` and `age`:

```{r}
lm.fit = lm(medv~lstat+age, data=Boston)
summary(lm.fit)
```

Fit a model using all 13 predictors:

```{r}
lm.fit = lm(medv~., data=Boston)
summary(lm.fit)
```

Fit a model using all predictors except for `age`:

```{r}
lm.fit.no.age = lm(medv~.-age, data=Boston)
summary(lm.fit.no.age)
```

# 3.6.4 Interaction Terms

Use the syntax `x1:x2` to add an interaction term between `x1` and `x2`, or use
`x1*x2` to include both predictors and the interaction term at the same time (a
shorthand for `x1+x2+x1:x2`):

```{r}
summary(lm(medv~lstat*age, data=Boston))
```

# 3.6.5 Non-linear Transformations of the Predictors

To create nonlinear transformations of predictors in the context of a linear
model formula, use the built-in `I()` function to ensure that `^` is treated
"as-is" (and not using the formula-specific syntax):

```{r}
lm.fit2 = lm(medv~lstat+I(lstat^2), data=Boston)
summary(lm.fit2)
```

We can use the `anova` function to test the hypothesis that the model with
a nonlinear term is superior to the original model (the null hypothesis is that
both models perform equally well):

```{r}
lm.fit = lm(medv~lstat, data=Boston)
anova(lm.fit, lm.fit2)
```

Plot the diagnostics for the new model to confirm that the residual plot has
improved:

```{r}
par(mfrow=c(2,2))
plot(lm.fit2)
```

For higher-order polynomials, use the built-in `poly()` function:

```{r}
lm.fit5 = lm(medv~poly(lstat, 5), data=Boston)
summary(lm.fit5)
```

# 3.6.6 Qualitative Predictors

Examine the predictors in the `Carseats` dataset:

```{r}
names(Carseats)
```

In this case, `ShelveLoc` is a qualitative predictor, which can take the value of
`Bad`, `Medium`, or `Good` (corresponding to the placement of the car seat on
the store shelf). Given such a qualitative variable, R generates dummy variables
automatically.

Fit a multiple regression model with some interaction terms:

```{r}
lm.fit = lm(Sales~.+Income:Advertising+Price:Age, data=Carseats)
summary(lm.fit)
```

Use the `constrasts()` function to read out the coding that R uses for the dummy
variables under the hood:

```{r}
contrasts(Carseats$ShelveLoc)
```

Since the coefficients for `ShelveLocGood` and `ShelveLocMedium` are both
positive and significant in the summary above, we can say with confidence that `Good`
and `Medium` shelf positions both predict higher sales than `Bad` shelf
positions. Further, since the coefficient for `ShelveLocGood` is greater than
the coefficient for `ShelveLocMedium`, we can be confident that a `Good` shelf
position will be better for sales than a `Medium` shelf position.

# 3.6.7 Writing Functions

Write a simple function to load the `MASS` and `ISLR` libraries:

```{r}
LoadLibraries = function() {
    library(ISLR)
    library(MASS)
    print("The MASS and ISLR libraries have been loaded.")
}
```

Call the function:

```{r}
LoadLibraries()
```
